### 2016-01-17 paper/language.tex
added syntax figure to paper

\section{Language}
\label{sec:language}

\subsection{Syntax}
Define the grammar of each of the languages - regular language, the macro language on top of it, and maybe the IR language.

% grammar
\newcommand{\BNFALT}{\;\;|\;\;}
\newcommand{\hdr}[2]{\flushleft \chdr{#1}{#2}}
\newcommand{\chdr}[2]{\textbf{#1} {#2} \\ \centering}

\begin{figure*}
  \begin{minipage}[t]{.42\linewidth}
  \hdr{\large Syntax}{}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{rclr}
    \hline
     pol &::=& c_1, \dots, c_n & \textit{constraints} \\
     c   &::=& p \Rightarrow r_1 \gg \dots \gg r_m & \textit{preferences} \\
     p   &::=& true & \textit{true} \\
      &\BNFALT& \neg p & \textit{negation} \\
      &\BNFALT& p_1 \vee p_2 & \textit{disjunction} \\
      &\BNFALT& p_1 \wedge p_2 & \textit{conjunction} \\
      &\BNFALT& prefix = x & \textit{prefix test} \\
      &\BNFALT& comm = t & \textit{community test} \\
     r   &::=& n & \textit{AS number} \\
      &\BNFALT& in & \textit{internal loc} \\
      &\BNFALT& out & \textit{external loc} \\
      &\BNFALT& r_1 \cup r_2 & \textit{union} \\
      &\BNFALT& r_1 \cap r_2 & \textit{intersection} \\
      &\BNFALT& r_1 \cdot r_2 & \textit{concatenation} \\
      &\BNFALT& !(r) & \textit{path negation} \\
      &\BNFALT& r^* & \textit{repetition} \\
  \end{array} \]

  \end{minipage}
  %
  ~~
  \vrule
  ~~
  %
  \begin{minipage}[t]{.5\linewidth}
  \hdr{\large Propane Macros}{}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{rcl}
    \hline
    only(X)       & = & X^* \\
    never(X)      & = & (!X)^* \\
    through(X)    & = & out^* \cdot in^* \cdot X \cdot in^* \cdot out^* \\
    upstream(X)   & = & out^* \cdot (X \cap out) \cdot out^* \cdot in^+ \cdot out^* \\
    downstream(X) & = & out^* \cdot in^+ \cdot out^* \cdot (X \cap out) \cdot out^* \\
    end(X)        & = & (out^* \cdot in^* \cdot (X \cap in)) \cup \\
                  & = & (out^* \cdot in^+ \cdot out^* \cdot (X \cap out)) \\
    start(X)      & = & ((X \cap in) \cdot in^* \cdot out^*) \cup \\
                  & = & ((X \cap out) \cdot out^* \cdot in^+ \cdot out^*) \\
    exit(X)       & = & (out^* \cdot in^* \cdot (X \cap in) \cdot out \cdot out^*) \cup \\
                  &        & (out^* \cdot in^+ \cdot (X \cap out) \cdot out^*) \\
    enter(X)      & = & (out^* \cdot out \cdot (X \cap in) \cdot in^* \cdot out^*) \cup \\
                  &        & (out^* \cdot (X \cap out) \cdot in^+ \cdot out^*) \\
    link(X,Y)     & = & (out^* \cdot in^* \cdot X \cdot Y \cdot in^* \cdot out^*) \cup \\
                  & = & (out^* \cdot (X \cap out) \cdot (Y \cap in) \cdot in^* \cdot out^*) \cup \\
                  & = & (out^* \cdot in^* \cdot (X \cap in) \cdot (Y \cap out) \cdot out^*) \\
  \end{array} \]

  \end{minipage}

  \hrulefill

  \caption{Regular Intemediate Language (RIL) syntax (left), and 
           Propane language expansions (right).}
  \label{fig:semantics}
\end{figure*}

\subsection{Semantics}
Give the semantics in terms of ranked paths.
Extension of regular expression semantics with the
additional constraint that we take only valid topology paths and that their are no paths with
repeated locations in the path. One tricky point is that we care about paths starting or ending at locations that are
connected to end-hosts in some way. And we want the length of the path to be a secondary tie-breaker.

\subsection{Theory}
What are the properties of this language? Ideally, a proof here about its expressiveness. We can perhaps do everything
BGP can do? We can't do shortest paths routing with BGP, but we could with OSPF though? Clearly we can only describe
policies that are stable under all possible failure scenarios (at least internally stable, could still cause) routing
loops with interactions with neighbors.