### 2016-01-22 paper/paper.tex
Memoize simulation relation.

\documentclass{sig-alternate-10pt}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{times}
\usepackage{xspace}
\usepackage{listings}

\usepackage{tikz}
\usetikzlibrary{arrows,automata,positioning}

\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\newcommand{\ratul}[1]{\textcolor{blue}{[ratul: #1]}}
\newcommand{\ryan}[1]{\textcolor{green}{[ryan: #1]}}

\newcommand{\sysname}{{\small \sf Propane}\xspace}

\newcommand{\para}[1]{\paragraph*{\textbf{#1}}}

\newcommand{\set}[1]{\ensuremath{\{ #1 \} }}
\newcommand{\abs}[1]{\ensuremath{ \lvert #1 \rvert }}
\newcommand{\path}[2]{ #1 \mapsto \ensuremath{#2} }

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\title{Toward Centrally Programming Distributed Control Planes}
\author{Paper \#XXX}

\maketitle

%\category{CR-number}{subcategory}{third-level}
%% general terms are not compulsory anymore,
%% you may leave them out
%\terms
%term1, term2
%\keywords
%keyword1, keyword2
\input{abstract}
\input{introduction}
\input{background}
\input{motivation}
\input{propane}
\input{compilation}
\input{evaluation}
\input{discussion}
\input{related}
\input{conclusions}

\bibliographystyle{abbrv}
\bibliography{references}

% The bibliography should be embedded for final submission.

\end{document}