### 2016-01-22 paper/expressiveness.tex
Adding abstract.

\para{Expressiveness}

\todo{Do we need all of this}