### 2016-01-22 paper/expressiveness.tex
Adding compiled policy figures.

\section{Expressiveness}

In this section we investigate correctness and expressiveness properties of \sysname. We are mainly concerned with the following questions: (1) does the distributed, compiled policy faithfully implement the user's policy regardless of failures? (2) Are the resulting BGP configurations stable?, and (3) what polcies are or are not expressible in \sysname?

In this section, we compare the expressiveness of the \sysname language to that of BGP.

\todo{Do we need all of this}