### 2016-01-23 paper/related.tex
related work

\section{Related Work}

\begin{itemize}
	\item Regular path queries
	\item Lots of background on BGP
	\item Feamster rcc stuff
	\item Logic programming BGP configs
\end{itemize}

\dpw{We could also consider citing Boon's stuff on declarative routing.}

\paragraph*{SDN Programming Languages}
The design of \sysname{} was heavily influenced by recent SDN programming
languages such as Merlin~\cite{foster:merlin}, FatTire~\cite{fattire},
NetKAT~\cite{netkat}, as well as Narayana's path queries~\cite{queries}.
Each of these languages is oriented around regular expressions, which
describe paths through a network, and predicates, which classify packets.
In particular, FatTire allows programmers to define sets of paths together
with a fault tolerance level (\emph{i.e.,} tolerate 1 or 2 faults)
and the compiler will generate
OpenFlow 1.3 rules that satisfy the policy.  \sysname is somewhat more
expressive than FatTire as it allows users to specify preferences among
routes and it automatically generates distributed
implementations that will tolerate
any number of faults.  Because FatTire generates data plane rules up front,
before faults occur, specifying higher levels of fault tolerance comes 
at the cost of generating additional rules, which must fit in switch
memory.  In contrast, \sysname uses traditional distributed
control plane mechanisms to react to faults as they occur and to the
extent that they occur without imposing up front costs that may
necessitate overprovisioning.
Because of the differences in the underlying technology, the analyses
and compilation algorithms used in \sysname are quite different from
previous work on SDN.  Finally, in addition to using path-based abstractions
for intra-domain routing, \sysname uses them for inter-domain routing as
well, unlike any existing SDN language.

\paragraph*{Configuration Analysis}  The notion that
configuring network devices is difficult and error-prone is not new.  Past 
researchers have
tried to tackle this problem by analyzing existing 
firewall configurations~\cite{fang,lumeta,margrave} and
BGP configurations~\cite{feamster+:rcc,feamster:thesis,ipassure,batfish,bagpipe} and reporting errors or
inconsistencies when they are detected. 
Our research is complementary to these analysis
efforts.  Propane hopes to eliminate bugs by using higher-level
languages and a ``correct-by-construction''
methodology.  By proposing network administrators write configurations
at a high-level of abstraction, a whole host of low-level errors can be 
avoided.  We also believe the high-level, centralized nature of propane policies
has the benefit of making configurations
easier to understand and maintain.

\paragraph*{Configuration Synthesis}  
ConfigAssure~\cite{narain:lisa05,narain+:configassure} 
is another system designed to
help users define and debug low-level router
configurations.  Inputs to
ConfigAssure include a \emph{configuration database}, which contains a
collection of tuples over constants and configuration variables, and a
\emph{requirement}, which is a set of constraints.  For instance, the
tuple \texttt{hsrp(rexa,rA1,int(1),int(2))} states that interface
\texttt{rA1} on device \texttt{rexa} belongs to the HSRP group defined
by configuration variable \texttt{int(1)} with virtual IP address
defined by configuration variable \texttt{int(2)}.  The constraint
\texttt{hsrp\_subnet([rexa-rA1,rexb-rB1])} states that a database
contains tuples defining HSRP group identifiers and virtual IPs for
interfaces \texttt{rA1} and \texttt{rB1} on devices \texttt{rexa} and
\texttt{rexb}.  The authors use a combination of logic programming and
SAT solving to find concrete values for configuration variables such
as \texttt{int(1)}.  ConfigAssure has very broad scope, handling
configuration for a wide range of different protocols and many
different concerns.  In contrast, the scope of \sysname is much
narrower.  In return, \sysname offers compact, higher-level
abstractions customized for our domain, such as regular paths, as well
as domain-specific analyses customized to those abstractions, such as
our failure safety analysis.  The implementation technology is also
entirely different, as we define algorithms over automata and graphs
as opposed to using logic programming and SAT-based model-finding.