### 2016-01-27 paper/compilation.tex
edits to 5

\section{Compilation}
\label{sec:compilation}

The front-end (FE) of \sysname simplifies operators' task of describing preferred paths. That simplicity, however, comes at the cost of compilation complexity. The compiler must efficiently compute the sets of paths represented by the intersection of preferences and topology, compute which ones can be honored under failures, and ensure policy compliance under all possible failure cases.

%\subsection{Overview}

\begin{figure}[t!]
\centering
\includegraphics[width=.75\columnwidth]{figures/pipeline}
\caption{Compilation pipeline stages for Propane.}
\label{fig:pipeline}
\end{figure}

%In this section we briefly describe our compilation strategy for \sysname.

We handle this challenge by breaking compilation into multiple stages and developing efficient algorithms for each.
Figure~\ref{fig:pipeline} shows the 5 stage compilation pipeline to translate \sysname FE policies to device-local BGP policies. The first stage of the pipeline involves simple rewriting rules and substitutions from the FE to the core Regular Intermediate Representation (RIR). Policies in RIR are checked for well-formedness (e.g., never constraining traffic that does not enter the network), before being combined with topology to obtain the Product Graph Intermediate Representation (PGIR). The PGIR is a data representation that compactly captures the flow of BGP announcements subject to the policy and topology restrictions. We develop efficient algorithms that operate over the PGIR to ensure policy compliance under failures, avoid BGP instability, and prevent aggregation-induced black holes. Once the compiler determines that the PGIR is safe, it translates it to an abstract BGP (ABGP) representation. To make configurations more readable for human operators, and to reduce the size of the resulting configurations, the \sysname compiler makes several passes over the ABGP form. Finally, vendor-specific adaptors can translate from ABGP to concrete device configurations.

\subsection{Regular IR (RIR)}
\label{sec:rir}

The \sysname FE is just a thin layer atop a core, regular-expression language for describing preference-based path constraints. Regular expressions are an expressive formalism that have been used extensively for describing paths through graphs~\cite{bib:todo}, with applications to networks~\cite{bib:todo}. Our regular expression language differs from others by allowing operators to describe preferences between paths.

% grammar
\newcommand{\BNFALT}{\;\;|\;\;}
\newcommand{\hdr}[2]{\flushleft \chdr{#1}{#2}}
\newcommand{\chdr}[2]{\textbf{#1} {#2} \\ \centering}

\begin{figure*}
  \begin{minipage}[t]{.45\linewidth}
  \hdr{\large Syntax}{}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{rclr}
    \hline

     pol     &::=& p_1, \dots, p_n & \textit{constraints} \\
     p       &::=& t \mapsto r_1 \gg \dots \gg r_m & \textit{preferences} \\
     t       &::=& true & \textit{true} \\
         &\BNFALT& \neg t & \textit{negation} \\
         &\BNFALT& t_1 \vee t_2 & \textit{disjunction} \\
         &\BNFALT& t_1 \wedge t_2 & \textit{conjunction} \\
         &\BNFALT& prefix = x & \textit{prefix test} \\
         &\BNFALT& comm = c & \textit{community test} \\
     r       &::=& n & \textit{AS number} \\
         &\BNFALT& in & \textit{internal loc} \\
         &\BNFALT& out & \textit{external loc} \\
         &\BNFALT& r_1 \cup r_2 & \textit{union} \\
         &\BNFALT& r_1 \cap r_2 & \textit{intersection} \\
         &\BNFALT& r_1 \cdot r_2 & \textit{concatenation} \\
         &\BNFALT& !(r) & \textit{path negation} \\
         &\BNFALT& r^* & \textit{repetition} \\
     l       &::=& r_1 \rightarrow r_2 & \textit{link pairs} \\
     cc     &::=& agg(x, l) \BNFALT tag(c, t, l) & \textit{control constraints} \\
  \end{array} \]

  \end{minipage}
  %
  ~~
  \vrule
  ~~
  %
  \begin{minipage}[t]{.5\linewidth}
  \hdr{\large Propane Expansions}{}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{rcl}
    \hline
    any           & = & out^* \cdot in^+ \cdot out^* \\
    internal      & = & in^+ \\
    only(X)       & = & any \cap X^* \\
    never(X)      & = & any \cap (!X)^* \\
    through(X)    & = & out^* \cdot in^* \cdot X \cdot in^* \cdot out^* \\
    later(X)      & = & out^* \cdot (X \cap out) \cdot out^* \cdot in^+ \cdot out^* \\
    before(X)     & = & out^* \cdot in^+ \cdot out^* \cdot (X \cap out) \cdot out^* \\
    end(X)        & = & any \cap (\Sigma^* \cdot X) \\
    start(X)      & = & any \cap (X \cdot \Sigma^*) \\
    exit(X)       & = & (out^* \cdot in^* \cdot (X \cap in) \cdot out \cdot out^*) \cup \\
                  &        & (out^* \cdot in^+ \cdot (X \cap out) \cdot out^*) \\
    enter(X)      & = & (out^* \cdot out \cdot (X \cap in) \cdot in^* \cdot out^*) \cup \\
                  &        & (out^* \cdot (X \cap out) \cdot in^+ \cdot out^*) \\
    link(X,Y)     & = & any \cap (\Sigma^* \cdot X \cdot Y \cdot \Sigma^*) \\
    path(\vec{X}) & = & any \cap (\Sigma^* \cdot X_1 \dots X_n \cdot \Sigma^*) \\
    novalley(\vec{X}) & = & any ~ \cap \\
                  &   & !path(X_2,X_1,X_2) ~ \cap \dots \cap \\
                  &   & !path(X_n,X_{n-1},X_n) \\
  \end{array} \]

  \end{minipage}

  \hrulefill

  \caption{Regular Intermediate Language (RIL) syntax (left), and
           Propane language expansions (right).}
  \label{fig:rir-syntax}
\end{figure*}

\para{Syntax}

Figure~\ref{fig:rir-syntax} shows the RIR syntax. A policy consists of one or more constraints, each of which consists of a test on the type of route, and a corresponding set of preferred regular paths. Regular paths are regular expressions where the base characters are abstract locations---either a router or an AS. Special \textit{in} and \textit{out} symbols refer to any internal or external location respectively.
For example, the constraint
$$prefix=74.3.28.0/24 \mapsto (as200 \cdot in^+) \gg (as100 \cdot in^+)$$
describes a more-preferred set of paths for traffic announced by a prefix no less specific than $74.125.28.0/24$, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $in^+$ stands for $in \cdot in^*$, at least one internal hop. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

\sysname also supports constraints purely on the control-plane behavior of BGP. For example, prefix aggregation, which should not affect routing behavior, is an important optimization to reduce routing table size and churn in practice. Aggregation, for example, from internal to external locations, is specified using the same regular syntax as before:
$$agg(128.17.0.0/16, in \rightarrow out)$$
where the expression $in \rightarrow out$ refers to control messages flowing from any internal to any external location.

We list only the route aggregation and community tagging constraints in Figure~\ref{fig:rir-syntax}, but we also support other constraints such as limiting the maximum number of routes allowed between ASes, or enabling BGP multipath.

\para{Semantics}

The semantics of RIR is in terms of ranked paths. Each preference-based regular path constraint (of the form $r_1 \gg \dots \gg r_j$) maps to a set of concrete paths in the network that match one of $r_i$. We denote a network path as a string of abstract locations (routers or external ASes) of the form: $n_1 n_2 \dots n_k$. A regular expression $r$ matches path $p$, if $p \in \mathcal{L}(r)$ and $p$ is a topologically-valid path. We denote the length of a path $p$ as $\abs{p}$. A path $p$ will have a rank:
$$(\min_i \set{ p \in \mathcal{L}(r_i) }, \abs{p})$$
where the rank is lexicographically ordered according to (1) the most preferred regular expression matched, and (2) as a tie breaker, the length of the path. Lower ranks indicate \emph{more} preferred paths. The semantics is to allow traffic to be sent along any of the most preferred paths for each pair of starting and ending locations that appear in some valid specified path.

The set of ranked paths depends on which paths are valid in the topology, and thus when failures occur, the most preferred routes change. The \sysname compiler ensures that generated configurations for a policy always achieve the most preferred path possible given the failures in the topology, using only distributed mechanisms.

\para{Propane FE to RIR}

The main differences between the \sysname FE and the RIR are: (1) FE allows the programmer to specify constraints separately, and combine them  modularly, (2) FE provides high-level names that abstract sets of routes, and (3) FE allows the preference operator to be used locally.

A key constraint when translating FE to RIR is to ensure that all specified routes are well-formed. In particular, each regular path constraint $r$ must satisfy $r \subseteq out^* \cdot in^+ \cdot out^*$. This ensures that one can only talk about controlling traffic that goes through the user's network at some point.

The translation from \sysname to RIR is based on a set of rewriting rules.
The first step merges separate constraints. It takes the cross product of per-prefix constraints, where logical conjunction ($a \wedge b$) is replaced by intersection on regular constraints ($a \cap b$), logical disjunction is replaced by union, and logical negation ($\neg a$) is replaced by ($any \cap !(a)$), where $any$ ensures the routes are well-formed.
%
For example, in the data center configuration from \S\ref{sec:propane}, combining the \textit{Routing} and \textit{Ownership} policies results in the following RIR:

\begin{lstlisting}[mathescape=true]
$\path{p_{g1}}{any \cap end(A)}$
$\path{p_{g2}}{any \cap end(B)}$
$\path{p_{l1}}{\neg enter(out) \cap end(E)}$
$\path{p_{l2}}{\neg enter(out) \cap end(F)}$
$\path{true}{exit(out)}$
\end{lstlisting}

The next step rewrites the high-level constraints such as \textit{enter} according to the equivalences in Figure~\ref{fig:rir-syntax}. Since preferences can only occur at the outermost level for an RIR expression, the final step is to ``lift" occurrences of the preference operator in each regular expression. For example, the regular expression $a \cdot (b \gg c) \cdot d$ is lifted to $(a \cdot b \cdot d) \gg (a \cdot c \cdot d)$ by distributing the preference over the sequence operator. In general, we employ the follwing equivalence:
%
$$x \odot (y_1 \gg \dots \gg y_n) = x \odot y_1 \gg \dots \gg x \odot y_n$$
%
where $\odot$ stands for an arbitrary regular binary operator. Preferences nested under a unary operator, \textit{star} or \textit{negation}, are flagged by the compiler as invalid policies.

\subsection{Product Graph IR}

\newcommand{\state}[4]{\node[state,#3](#1)[#4]{#2};}
\newcommand{\transition}[4]{\path[->] (#1) edge [#4] node {#3} (#2);}

\begin{figure*}
  \begin{minipage}[t]{.5\linewidth}
  \hdr{\large Topology}{}
  \vspace*{-1\baselineskip}

  \includegraphics[width=.8\columnwidth]{figures/topology}

  \hdr{\large Policy Automata}{}
  \vspace*{1\baselineskip}

  \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=1.6cm]
    \state{0}{$0$}{              }{}
    \state{1}{$1$}{right of=0}{}
    \state{2}{$2$}{right of=1}{}
    \state{3}{$3$}{right of=2}{}
    \state{4}{$4$}{right of=3}{}
    \state{5}{$5$}{right of=4}{accepting}
    \transition{0}{1}{out}{}
    \transition{1}{2}{D}{}
    \transition{2}{3}{C}{}
    \transition{3}{4}{A}{}
    \transition{4}{5}{W}{}
  \end{tikzpicture}

  \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=1.6cm]
    \state{0}{$0$}{              }{}
    \state{1}{$1$}{right of=0}{}
    \state{2}{$2$}{right of=1}{}
    \state{3}{$3$}{right of=2}{}
    \state{4}{$4$}{right of=3}{accepting}
    \transition{0}{1}{out}{}
    \transition{1}{2}{in}{}
    \transition{2}{2}{A,C,D,E}{loop above}
    \transition{2}{3}{B}{}
    \transition{3}{3}{B}{loop above}
    \transition{3}{2}{A,C,D,E}{bend left}
    \transition{3}{4}{W}{}
  \end{tikzpicture}

  \end{minipage}
  %
  ~~
  ~~
  %
  \begin{minipage}[t]{.5\linewidth}
  \hdr{\large Product Graph}{}
  \vspace*{-1\baselineskip}
  %
  \includegraphics[width=.8\columnwidth]{figures/productgraph}
  \end{minipage}

  \hrulefill

  \caption{Example Product graph construction.}
  \label{fig:example-compilation}
\end{figure*}

Now that the user policy exists in a simplified form, we must consider topology. In particular, we want a compact representation that describes all the possible ways BGP route announcements can flow through the network subject to the policy and topology constraints. Our PGIR captures these dependencies by ``intersecting" each of the regular automata corresponding to the RIR path preferences, and the topology. Paths through the PGIR correspond to real paths through the topology that satisfy the user constraints.

%Finding paths through a graph subject to regular constraints has been studied extensively in the database literature~\ref{bib:todo}, and has been applied to networks in the past~\ref{bib:todo}.

\para{Formal definition}

% Formally define automata
The RIR policy is of the form $r_1 \gg \dots \gg r_j$. While paths talk about the direction traffic flows through the network, to implement the policy with BGP we are concerned about the way control-plane information is disseminated (i.e., route announcements flowing in the opposite direction). To capture this idea, for each regular expression $r_i$, we construct a deterministic finite state machine on the reversed regular expression. Each automaton is a tuple: ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ consists of all internal and external ASes, \ratul{ASes or locations?} $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
% Formally define topology
The topology is represented as a graph ($V, E$), which consists of a set of vertices $V$, and a set of directed edges $E \colon 2^{V \times V}$.
%
% Formally define product graph
The combined PGIR is a tuple: ($V'$, $E'$, $s$, $e$, $P$) with
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon 2^{V' \times V'}$,
a unique starting vertex $s$,
a unique ending vertex $e$,
and a preference function $P \colon V \rightarrow 2^{\set{1, \dots, j}}$ , which maps nodes in the product graph to a set of preference ranks.
For a vertex $v' = (v, \dots) \in V'$, we say $loc(v') = v$. We say a node $x$ shadows node $y$ in the product graph when $x \in V'$ and $y \in V'$ and $loc(x) = loc(y)$ but $x \neq y$. That is, $x$ shadows $y$ when they are different nodes in the PGIR that represent the same topology location.

\para{From RIR To PGIR}

% Formal product construction
Let $a_i$ and $b_i$ denote states in the regular policy automata.
The PGIR is constructed by adding an edge from $v_1 = (x, a_1, \dots, a_m)$ to $v_2 = (y, b_1, \dots, b_m)$ whenever $\sigma_i(a_i, y) = b_i$ for each $i$ and $(x,y) \in E$ is a valid topology link.
%
Additionally, we add edges from the start node $s$ to any $v = (x, a_1, \dots, a_m)$ when $\sigma_i(q_{0_i}, x) = a_i$ for each $i$.
%
The preference function $P$ for node $v = (x, a_1, \dots, a_m)$ is defined as $P(v) = \set{i~\vert~a_i \in F_i} $. That is, it records which preferences are achieved for the policy automata for each node in the product construction.
%
Finally, there is an edge from each node in the PGIR such that $P(v) \neq \emptyset$ to the special end node $e$.

Intuitively, PGIR tracks which states of each automaton the policy is in as route announcements move between routers.
%
Consider the topology in Figure~\ref{fig:example-compilation}. Suppose we want a primary route from W that enters the network from $A$, and utilizes the A--C and C--D links. We also want a backup route that enters from $B$, and utilizes the B--C link, but is otherwise unconstrained. For simplicity, we assume that the route can end in either $X$, $Y$, or $Z$. The RIR for the policy is:
%
$$(W \cdot A \cdot C \cdot D \cdot out) \gg (W \cdot B \cdot in^* \cdot out)$$
%
Figure~\ref{fig:example-compilation} shows the policy automata for each regular expression preference. Since we are interested in the flow of control messages, the automata match backwards.
%
The figure also shows the PGIR after intersecting the topology and policy automata. Every path in it corresponds to a concrete path in the topology. In particular, every path through the PGIR that ends at a node $v$ such that the preference function $P(v) = \set{i_1, \dots, i_m}$ is non-empty, is a valid topological path that satisfies the policy constraints and results in a particular path with preferences $i_1$ through $i_m$.
%
For example, the path $X \cdot D \cdot C \cdot A \cdot W$ is a valid path in the topology that BGP route announcements might take, which would lead to obtaining the most prefered preference $1$.
BGP control messages can start from peer X, which would match the $out$ transition from both automata, leading to state $1$ in the first automaton, and state $1$ in the second automaton. This is reflected in the product graph by the node with state $(1,1) X$. From here, if X were to advertise this route to D, it would result in the path $D X$, which would lead to state $2$ in the first automaton, and state $2$ in the second automaton, and so on. 
%
The ``--" state indicates it is dropped from the corresponding automaton. Since node $(5,-) W$ is in an accepting state for the first automata, it indicates this path has preference 1.

\para{Minimization}
\ratul{This minimization section can probably go entirely, shrunken, or moved to the implementation}
Before checking if policies captured by the PGIR are safe, we minimize the PGIR to better scale our failure safety checks. The minimization is based on the observation that although every path in the PGIR is a valid path in the topology, we do not want to consider loops when configuring the network. In particular, BGP's loop prevention mechanism, by which an AS rejects any route for which it is already in the AS path, means that loops should never occur.\footnote{ Misconfigured aggregation can still lead to loops in BGP.}
%
However, the PGIR may contain many paths that form loops in the topology. For example, in Figure~\ref{fig:example-compilation}, the path $W \cdot A \cdot C \cdot B \cdot W$ is a valid topological path that BGP announcements might take, leading to a path that satisfies the preference 1 policy, but which contains a loop.
%
Since we are primarily concerned with guaranteeing user policy correctness and safety with respect to network failures, it is useful to eliminate cases that are not possible due to loops. For example, if we can safely remove a PGIR edge, then we do not have to consider what happens when the edge has failed.

We can safely remove any PGIR node or edge as long as it never appears on any \emph{simple} path from the \textit{start} node to the \emph{end} node. For example, node $(1,1) W$ in Figure~\ref{fig:example-compilation} is never on a simple path to the end node since it most go through another $W$ in either case.
%
Although fully minimizing this graph is an NP-complete problem,  a simple and efficient algorithm based on graph dominators achieves largely the same effect, greatly simplifying the failure safety analysis.

A node $x$ dominates $y$ if $x$ appears on every path leading to $y$ from a source node. Many efficient algorithms exist for finding graph dominators~\ref{bib:todo}. For minimization, we compute the dominator set for each node in both the product graph $G$ with respect to the start node, and in the reversed product graph $G^R$ with respect to the end node. The following rules enable efficient minimization of the product graph.
%
\begin{itemize}
  \item Remove nodes that never reach the end node
  \item Remove nodes not reachable from the $start$ node
  \item Remove any node $x$ such that $x$ is dominated by $y$ in $G$ or $G^R$, and $loc(x) = loc(y)$
  \item Remove any edge from $x$ to $y$ in $G$ or $G^R$ if there is some node $z$ dominated by $y$ such that $loc(x) = loc(z)$
\end{itemize}
%
Repeated application of the above rules leads to removing the colored nodes and dashed lines in Figure~\ref{fig:example-compilation}.

\para{Failure Safety}

To implement path preferences in routing, BGP uses local preferences on a per-device basis. However, the distributed nature of BGP makes setting preferences locally to achieve a network-route routing policy difficult, particularly in the presence of failures. For example, imagine an extremely simple policy for the topology in Figure~\ref{fig:example-compilation}, which says to prefer one route over another:
%
$$(W \cdot A \cdot C \cdot D) \gg (W \cdot B \cdot C \cdot E)$$
%
How could such a policy be implemented in BGP? Suppose we simply set the local preference at router $C$ to prefer $D$ over $E$. However, now if the A -- C link fails, then suddenly $C$ has made the wrong decision and can not get either the primary or backup route specified, despite the fact that the $W \cdot B \cdot C \cdot E$ path is available.

At a first cut, any time a router must make a decision locally between several route options, there is the possibility that it might choose incorrectly. PGIR captures this notion of choice precisely. For example, router $D$ in Figure~\ref{fig:example-compilation} appears only once in the PGIR, possibly receiving announcements from $X$ or $Y$. However, regardless of whether $D$ chooses a route from $X$ or $Y$, the set of paths allowed by the policy after going through $D$ will be the same in either case. This remains true despite any failures that might have occurred in the network. Thus $D$ can safely prefer $X$ and $Y$ equally.

The more challenging case is when a topology location occurs in mutliple contexts in the PGIR. For example, in the compilation example from Figure~\ref{fig:example-compilation}, the topology node $C$ can receive an announcement from $E$ and later achieve the backup path, or it can receive an announcement from its neighbor $D$ and later achieve either the primary or backup path. Is it safe for $C$ to prefer its neighbor $D$ over its other neighbor $E$? The important observation is that, if $C$ prefers $D$, then it is never worse off---it will always achieve at least as good a path as if it had choosen $E$.
For example, suppose $C$ chooses a route from $D$, but cannot achieve its primary path because the $A$ -- $W$ link has failed. In this case, the advertisement from $C$ will still be sent along towards the ultimate backup location $W$. Since the $(3,2) C$ node has the same one-step and two-step next hops as node $(-,2) C$ in the product graph, no possible failure will prevent a route advertisment from reaching $(-,4) W$ that wouldn't have otherwise prevented it if $C$ had choosen $E$.

In general, ensuring that an individual router's preferences respect the policy under all possible failures is hard, and enumerating all failures is intractable. We thus adopt a conservative analysis based on the observations above. The high-level idea is to order each PGIR node according to a \textit{can prefer} relation. For example, node $(3,2) C$ can be preferred to node $(-,2) C$, but not the other way around. Intuitively, a node $N_1$ can be preferred to another node $N_2$ when, for each preference $N_2$ might achieve, there is a better preference that $N_1$ will achieve regardless of failures. Formally this means that $N_1$ can be preferred to $N_2$ when:
%
$$\forall i, \exists j, j \leq i \wedge protect(G_j, N_1, G_i, N_2)$$
%
where \textit{protect} means that, from $N_1$ on the PGIR restricted nodes that can potentially achieve preference $j$ or better, there are at least as many paths as from $N_2$ restricted to $G_i$

\begin{algorithm}[t!]
\caption{Failure Protection}
\label{alg:failures}
\begin{algorithmic}[1]
\Procedure{Protect($G_1$, $N_1$, $G_2$, $N_2$)}{}
  \If {$loc(N_1) \neq loc(N_2)$} \Return false
  \EndIf
  \State $q \gets Queue()$
  \State $q.Enqueue (N_1, N_2)$
  \While {$!q.Empty()$}
    \State $(n_1,n_2) \gets q.Dequeue()$
    \For {$x$ in $adj(G_2, n_2)$}
      \If {some $y$ in $adj(G_1,n_1)$ shadows $x$}
        \If {$(x,y)$ not marked}
          \State {mark $(x,y)$ as seen}
          \State $q.Enqueue(x,y)$
        \EndIf
      \Else { \Return $false$}
      \EndIf
    \EndFor
  \EndWhile
  \Return true
  \EndProcedure
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{alg:failures} defines what it means for one node to \textit{protect} against the failures of another. It walks from nodes $N_1$ and $N_2$ in $G_j$ and $G_i$ respectively, and ensures that for every \textit{step} $N_2$ can take to some new topology location, $N_1$ can, at the very least, take an equivalent step. Thus any paths leading from $N_2$ will have an equivalent path over the same topology locations from $N_1$. Algorithm~\ref{alg:failures} works because each PGIR node can have at most one outgoing neighbor with the same topological location, so the next hop neighbors can be checked without backtracking. The algorithm terminates since the number of related states $(x,y)$ that can be expolored is finite.
\ryan{Add the new part that goes up to a dominating predecessor when no other option}

Local preferences are now obtained by, for each router in the topology, sorting the corresponding PGIR nodes according to \textit{can prefer} relation. If two nodes are incomparable, then the compiler rejects the policy as unimplementable.
\ryan{We are checking if the preorder forms a partial order by satisfying asymmetry}

\para{Avoiding Simple Paths}

The checks for failure safety described above overlook a critical point: even if the analysis determines that a node $x$ protects against failures of another node $y$, it may be tricked into thinking $x$ has a particular path that will actually be discarded by BGP's loop prevention mechanism due to locations previously traversed before arriving at $x$. To avoid this situation, the compiler conservatively checks that $x$ protects against $y$ without using any nodes that shadow another node that appears above $x$ in the PGIR (i.e., nodes reachable from $x$ in $G_j^R$). It is fine however, to reuse the same nodes that appear above $x$, since any path that went through another node $z$ to get to $x$, but then looped through $z$ again, would be obtainable from $z$ itself.
\ryan{rework this}

\subsection{Abstract BGP}

%\begin{figure}[t!]
%\centering
%\includegraphics[width=.9\columnwidth]{figures/config}
%\caption{Abstract BGP configuration.}
%\label{fig:abgp-config}
%\end{figure}

\newcommand{\highlight}[1]{%
  \colorbox{red!50}{$\displaystyle#1$}}
\newcommand{\Router}[1]{ \textbf{Router #1:} }
\newcommand{\REGEX}[1]{ \text{regex}(#1) }
\newcommand{\PEER}{ \text{peer} }
\newcommand{\COMM} {\text{comm}}
\newcommand{\MED} {\text{MED}}

\begin{figure}[t!]
\begin{lstlisting}[frame=single, mathescape=true]
$\Router{A}$
  Match $\PEER=C, \COMM=(3,2)$
    Export $\COMM \leftarrow (4,2), \MED \leftarrow 80, \PEER \leftarrow W$
$\Router{B}$
  Match $\PEER = C, \COMM = (-,2)$
    Export $\COMM \leftarrow (-,3), \COMM \leftarrow \text{noexport},$
           $\MED \leftarrow 81, \PEER \leftarrow W$
  Match $\PEER = C, \COMM = (3,2)$
    Export $\COMM \leftarrow (-,3), \COMM \leftarrow \text{noexport},$
           $\MED \leftarrow 81, \PEER \leftarrow W$
$\Router{C}$-
  Match$[LP=99]$ $\PEER = E, \COMM = (-,2) $
    Export $\COMM \leftarrow (-,2), \PEER \leftarrow B$
  Match $\PEER = D, \COMM = (2,2)$
    Export $\COMM \leftarrow (3,2), \PEER \leftarrow A,B$
$\Router{D}$
  Match $\REGEX{X + Y}$
    Export $\COMM \leftarrow (2,2), \PEER \leftarrow C$
$\Router{E}$
  Match $\REGEX{Z}$
    Export $\COMM \leftarrow (-,2), \PEER \leftarrow C$
\end{lstlisting}
\label{fig:config}
\caption{Abstract BGP router configurations.}
\end{figure}

%\begin{figure}[t!]
%\begin{lstlisting}[frame=single, mathescape=true]
%$\Router{A}$
%  Match $\PEER=C$
%    Export $\MED \leftarrow 80, \PEER \leftarrow W$
%$\Router{B}$
%  Match $\PEER = C$
%    Export $\COMM \leftarrow \text{noexport}, \MED \leftarrow 81, \PEER \leftarrow W$
%$\Router{C}$
%  Match$[LP=99]$ $\PEER = E $
%    Export $\PEER \leftarrow B$
%  Match $\PEER = D$
%    Export $\PEER \leftarrow A,B$
%$\Router{D}$
%  Match $\REGEX{X + Y}$
%    Export $\PEER \leftarrow C$
%$\Router{E}$
%  Match $\REGEX{Z}$
%    Export $\PEER \leftarrow C$
%\end{lstlisting}
%\label{fig:config-min}
%\caption{Abstract BGP minimized configurations}
%\end{figure}

The final stages of compilation consist in translating policies from PGIR to a simple, vendor-neutral abstraction of BGP (ABGP), and then translating from ABGP to actual device configurations.

\para{From PGIR to ABGP}

Having found a total ordering on node preferences in the product graph from the failure safety analysis, the translation to ABGP becomes straightforward. The idea is to encode the state of the automaton using BGP community values. Each router will match based on its peer and a community value corresponding to the state of the product graph, and then update the state before exporting to the neighbors permitted according to the product graph. For example, router $A$ from the example in Figure~\ref{fig:example-compilation} will allow an advertisement from its peer $C$ with a community value for state $(3,2)$ (and deny anything else). If it sees such an advertisement, it will remove the old community value, and add a new community value for state $(4,2)$ before exporting the route to its neighbor $W$.

To ensure preferred paths are always obtained, for each router $r$ in the topology, the compiler will simply set a higher local preference for neighbors of a more-preferred node for $r$ in the product graph. For example, $C$ will prefer an advertisement from $D$ in state $(2,2)$ over an advertisement from $E$ in state $(-,2)$.

Since the compiler is only able to control community tagging only for routers under the control of the AS being programmed, it cannot match on communities for external ASes. Instead, it simply translates matches from external ASes into a BGP regular expression filter. For example, node $D$ in Figure~\ref{fig:example-compilation} would match the single hop external paths $X$ or $Y$ and prefer them equally. In general, if routes are allowed from beyond $X$ or $Y$, these will also be captured in the BGP regular expression filters. In practice, we model the unknown AS topology as a special node in the product graph that generates a filter to match any sequence of ASes.

Finally, in our example, the external AS $W$ would need to prefer our internal router $A$ over $B$. In general, it is not possible to control traffic entering the network beyond certain special cases. In this example however, the BGP MED attribute can influence $W$ to prefer $A$ over $B$ since the preference is for a single external peer.  Additionally, the compiler can use the BGP no-export community can ensure that no other AS beyond $W$ can send us traffic. The compiler can perform a simple analysis to determine when it can utilize various BGP special attributes to ensure traffic enters the network in a particular way by looking at links in the product graph that cross from the internal topology to the external topology. Aggregation is another tool commonly used to influence how traffic can enter the network, by relying on BGP's longest prefix matching default. Since aggregates are provided as additional user constraints, the compiler can also make use of this information to check that incoming preferences can be met given the aggregates specified (e.g., to ensure that all external ASes prefer to enter from one location over another).

Figure~\ref{fig:config} shows the full configuration from the compilation example. To make configurations more compact and readable, the \sysname compiler will rewrite configurations in ABGP - for example removing community tagging and matching when there is no ambiguity. Figure~\ref{config-min} shows the minimized router configurations.

\subsection{Aggregation-Induced Black Holes}

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/aggregation}
\caption{Aggregation safety.}
\label{fig:aggregation-safety}
\end{figure}

As was demonstrated in Example 2 of Section~\ref{sec:motivation}, the use of aggregation in BGP can lead to subtle black-holing traffic when failures occur. Deciding when this can happen requires knowledge of the routing policy and not just the topology. For instance, a policy might require all traffic for a particular prefix going through a router where an aggregate is advertised to go over a single link, even if there are several available in the toplogy. If the single link fails, then a black hole might be introduced. Fortunately, the PGIR captures this information precisely, describing the how information meeting the policy flows over the topology.

We can frame the aggregation problem as a problem of connectivity in the PGIR. More specifically, for each prefix that falls under an aggregate, we find a lower bound on the number of failures that would disconnect where the prefix originates from where its more specific aggregate is located. The difficulty lies in the fact that the same links in the topology can appear in multiple places in the PGIR. We adopt the following simple strategy to lower bound the number of failures:
\begin{enumerate}
	\item Pick a random path through the product graph from start to end.
	\item Remove all edges in the product graph corresponding to the set of topological edges choosen.
	\item Repeat until no path exists.
\end{enumerate}
Since each path through the PGIR will result in a new edge-disjoint path through the topology (since we have removed all already used edges), and since the number of edge-disjoint paths through the topology is no more than the min-cut. \todo{clarify this}

For example, imagine we use the earlier data center example from Section~\ref{sec:motivation}, with the policy: $p_{g1} \mapsto end(A)$, where $p_{g1}$ falls under the $p_{g^*}$ aggregate. The data center topology and a simplification of the corresponding product graph are shown in Figure~\ref{fig:aggregation-safety}. Since the compiler knows an aggregate will be placed at $X$, and it knows that, for prefix $p_{g1}$ the route will originate at $A$, we can compute the number of failures it would take to disconnect $A$ from $X$ in the product graph. In this case, we could remove the $A$ -- $D$ -- $X$ path first. We would then need to remove any other $A$ -- $D$ or $D$ -- $X$ links from the product graph, though in this case there are none. Next, we could remove the links along the $A$ -- $C$ -- $X$ path, repeating the process. Because afterwards $A$ is disconnected from $X$, the compiler knowns that 2 is a lower bound on the number of failures for aggregation safety for prefix $p_{g1}$. This process is then repeated for other aggregation locations (e.g., $Y$).